\pkg{Rgbp}: An R Package for Hierarchical Modeling and Method Checking

### 2013-06-10 jss/article.tex
finally up to 8 schools example. Will do tomorrow morning
8 schools disc

\subsection{Normal: 8 Schools}
The Education Testing Service (ETS) conducted randomized experiments in eight separate schools (group) to test whether students (unit) SAT scores are effected by coaching. The dataset contains the estimated coaching effects on SAT scores ($y_{j}, j=1, \ldots, 8$) and standard errors ($se_{j}, j=1, \ldots, 8$) of the eight schools \cite{1981}.
\begin{CodeChunk}
\begin{CodeInput}
R> y  <- c(12, -3, 28,  7,  1,  8, 18, -1)
R> se <- c(18, 16, 15, 11, 11, 10, 10,  9)
\end{CodeInput}
\end{CodeChunk}

@@

%TODO: Resume editing from here
\section[Method Checking]{Method Checking}
Like the two sides of the same coin, checking a statistical model always comes with fitting a model. If a fitted model cannot pass a validation or checking process, we usually go back and forth from estimation and checking steps iteratively. In this sense, checking a fitted model is an interactive procedure for the model justification.

@@

\section[Examples]{Examples}
\subsection[Known Second-level Mean]{31 Hospitals: Known Second-level Mean}
Suppose you live in the New York state (NY) and have been suffering from severe coronary heart disease (hopefully not). If you are supposed to receive the coronary artery bypass graft (CABG) surgery soon, you might want to find the most famous hospital for dealing with such a surgery. On top of that, if you can figure out each hospital's ability to handle this surgery, it will be useful for your decision in choosing a hospital.

@@

A function ``\code{summary}'' shows selective information on hospitals and more detailed estimation result as below. To be specific, it displays some hospitals (not all as above) with minimum, median, and maximum caseloads ($n_{j}$). On top of that, more specific estimation results, such as the estimation result of $\alpha\equiv\log(1/r)$, follow. Note that when we do not know the prior mean in advance unlike this hospital problem, \code{gbp} fits a linear regression for the Normal model, a log-linear regression for the Poisson model, or a logistic regression for the Binomial model and a summary of regression fit will appear.
\begin{CodeChunk}
\begin{CodeInput}
R> summary(p)
\end{CodeInput}
\begin{CodeOutput}
Main summary:

@@

\subsection[Unknown Second-level Mean and No Covariate]{8 Schools: Unknown Second-level Mean and No Covariate}
\begin{CodeChunk}
\begin{CodeInput}
R> g <- gbp(y, se, model = "gr")
R> g
\end{CodeInput}
\begin{CodeOutput}
Summary for each unit (sorted by n):

         obs.mean   se prior.mean shrinkage low.intv post.mean upp.intv post.sd
5           -1.00  9.0      8.168     0.408  -13.297     2.737   16.692   7.634
2            8.00 10.0      8.168     0.459   -7.255     8.077   23.361   7.810
7           18.00 10.0      8.168     0.459   -1.289    13.484   30.821   8.176
4            7.00 11.0      8.168     0.507   -8.780     7.592   23.602   8.257
6            1.00 11.0      8.168     0.507  -13.027     4.633   20.131   8.441
1           28.00 15.0      8.168     0.657   -2.315    14.979   38.763  10.560
3           -3.00 16.0      8.168     0.685  -17.130     4.650   22.477  10.096
8           12.00 18.0      8.168     0.734  -10.208     9.189   29.939  10.227
colMeans     8.75 12.5      8.168     0.552   -9.163     8.168   25.723   8.900
\end{CodeOutput}
\end{CodeChunk}

@@

%\begin{figure}[h]
%\begin{center}
%\includegraphics[scale=0.3]{school1.png}
%\caption{Shrinkage and interval plots of 8 schools}
%\end{center}
%\end{figure}

%\begin{CodeChunk}
%\begin{CodeInput}
%R> gcv <- coverage(g, nsim = 10000)
%\end{CodeInput}
%\end{CodeChunk}
%\begin{figure}[h]
%\begin{center}
%\includegraphics[width = 5.5cm]{school2.png}
%\end{center}
%\end{figure}

>>>>

\subsection{Normal: 8 Schools} \label{8schooldata}
The Education Testing Service (ETS) conducted randomized experiments in eight separate schools (group) to test whether students (unit) SAT scores are effected by coaching. The dataset contains the estimated coaching effects on SAT scores ($y_{j}, j=1, \ldots, 8$) and standard errors ($se_{j}, j=1, \ldots, 8$) of the eight schools \cite{1981}.
\begin{CodeChunk}
\begin{CodeInput}
R> y  <- c(12, -3, 28,  7,  1,  8, 18, -1)
R> se <- c(18, 16, 15, 11, 11, 10, 10,  9)
\end{CodeInput}
\end{CodeChunk}

@@

%TODO: Haven't proof read method checking section yet
\section[Method Checking]{Method Checking}
Like the two sides of the same coin, checking a statistical model always comes with fitting a model. If a fitted model cannot pass a validation or checking process, we usually go back and forth from estimation and checking steps iteratively. In this sense, checking a fitted model is an interactive procedure for the model justification.

@@

%TODO: Proof read Hospital and Baseball examples
\section[Examples]{Examples}
\subsection[Known Second-level Mean]{31 Hospitals: Known Second-level Mean}
Suppose you live in the New York state (NY) and have been suffering from severe coronary heart disease (hopefully not). If you are supposed to receive the coronary artery bypass graft (CABG) surgery soon, you might want to find the most famous hospital for dealing with such a surgery. On top of that, if you can figure out each hospital's ability to handle this surgery, it will be useful for your decision in choosing a hospital.

@@

A function ``\code{summary}'' shows selective information on hospitals and more detailed estimation result as below. To be specific, it displays some hospitals (not all as above) with minimum, median, and maximum caseloads ($n_{j}$). On top of that, more specific estimation results, such as the estimation result of $\alpha\equiv\log(1/r)$, follow. Note that when we do not know the prior mean in advance unlike this hospital problem, \code{gbp} fits a linear regression for the Normal model, a log-linear regression for the Poisson model, or a logistic regression for the Binomial model and a summary of regression fit will appear.
\begin{CodeChunk}
\begin{CodeInput}
R> summary(p)
\end{CodeInput}error:%20LaTeX%20Error:%20File%20`a4wide.sty'%20not%20found.
\begin{CodeOutput}
Main summary:

@@

\subsection[Unknown Second-level Mean and No Covariate]{8 Schools: Unknown Second-level Mean and No Covariate}
As stated in \ref{8schooldata} the Education Testing Service (ETS) conducted randomized experiments in eight separate schools (group) to test whether students (unit) SAT scores are effected by coaching. The dataset contains the estimated coaching effects on SAT scores ($y_{j}, j=1, \ldots, 8$) and standard errors ($se_{j}, j=1, \ldots, 8$) of the eight schools \cite{1981}.
\begin{CodeChunk}
\begin{CodeInput}
R> y  <- c(12, -3, 28,  7,  1,  8, 18, -1)
R> se <- c(18, 16, 15, 11, 11, 10, 10,  9)
\end{CodeInput}
\end{CodeChunk}

Due to the nature of the test each school's coaching effect has an approximately Normal sampling distribution with known sampling variance, \emph{i.e.}, standard error of each school is assumed to be known or to be accurately estimated. Hence, we can use \code{gbp} to fit this hierarchical model:

\begin{CodeChunk}
\begin{CodeInput}
R> data(schools)
R> attach(schools)
R> g <- gbp(y, se, model = "gr")
R> g
\end{CodeInput}
\begin{CodeOutput}
Summary for each unit (sorted by n):

         obs.mean   se prior.mean shrinkage low.intv post.mean upp.intv post.sd
5           -1.00  9.0      8.168     0.408  -13.297     2.737   16.692   7.634
2            8.00 10.0      8.168     0.459   -7.255     8.077   23.361   7.810
7           18.00 10.0      8.168     0.459   -1.289    13.484   30.821   8.176
4            7.00 11.0      8.168     0.507   -8.780     7.592   23.602   8.257
6            1.00 11.0      8.168     0.507  -13.027     4.633   20.131   8.441
1           28.00 15.0      8.168     0.657   -2.315    14.979   38.763  10.560
3           -3.00 16.0      8.168     0.685  -17.130     4.650   22.477  10.096
8           12.00 18.0      8.168     0.734  -10.208     9.189   29.939  10.227
colMeans     8.75 12.5      8.168     0.552   -9.163     8.168   25.723   8.900
\end{CodeOutput}
\end{CodeChunk}
The output from \code{gbp} provides an easy to read summary of the results of the estimation. Here we see that the posterior mean estimates are simply a linear combination

\begin{CodeChunk}
\begin{CodeInput}
R> plot(g)
\end{CodeInput}
\end{CodeChunk}

\begin{figure}[h]
\begin{center}
\includegraphics[scale=0.3]{school1.png}
\caption{Shrinkage plot and 95\% interval plot for 8 schools}
\end{center}
\end{figure}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> gcv <- coverage(g, nsim = 10000)
\end{CodeInput}
\end{CodeChunk}
\begin{figure}[h]
\begin{center}
\includegraphics[width = 5.5cm]{school2.png}
\caption{Coverage plot via method checking for 8 schools}
\end{center}
\end{figure}
